Phylogenetics and the human microbiome.

### 2014-04-17 sbreview.tex
Moving from plainnat_abbrv to sysbio.
Loads of bib fixes.

The name UniFrac is a contraction of ``Unique Fraction,'' which refers to the fact that the original UniFrac definition compares the fraction of edges that connect only tips from one sample via the shortest path rather than edges that connect between samples \citep[Fig.~\ref{fig:unifrac};][]{LozuponeKnightUniFrac05}.
Weighted UniFrac is an abundance weighted version \citep{LozuponeEaWeightedUnifrac07}.
These dissimilarity measures have hundreds of citations.
\citet{evans2012phylogenetic} showed that weighted UniFrac is in fact a specific case of the earth-movers distance, and that the commonly-used randomization procedure for significance estimation has a central limit theorem approximation.
The earth-movers distance between two probability distributions in this case can be defined using a physical analogy as the minimum amount of ``work'' required to move piles of dirt (each dirt pile configuration sized in proportion to one probability distribution) in one configuration to another along the tree, where work is defined as mass times distance.
In this case the size of the dirt piles is proportional to the number of reads mapping to that location in the tree (Fig.~\ref{fig:dirtpiles}).
\citet{chen2012associating} have shown that a partially abundance-weighted variant of UniFrac may have greater power to resolve community differences than either unweighted or weighted UniFrac.
There are clear connections between UniFrac and phylogenetic diversity explored by \cite{Faith2009-gs}, who point out related measures in Faith's earlier work.

@@

There are also other methods with the less ambitious goal of mapping sequences of unknown origin into a so-called fixed \textit{reference tree}, sometimes with uncertainty estimates.
These programs \citep{wu2008simple,monierEaLargeViruses08,vonMeringEaQuantitative08,stark2010mltreemap,matsen2010pplacer,berger2011performance} have various speeds and features.
This work has also spurred development of specialized alignment tools for this mapping process.
\citet{berger2011aligning} focus on the problem of inferring the optimal alignment and insertion of sequences into a tree.
\citet{mirarabsepp} use data set partitioning to improve alignments on subsets of taxa specifically for this application.

@@

\noindent
Figure 1: \pdLegend

\noindent
Figure 2: \unifracLegend

\noindent
Figure 3: \dirtpilesLegend

@@

\bibliographystyle{plainnat_abbrv}
\bibliography{sbreview}

>>>>

The name UniFrac is a contraction of ``Unique Fraction,'' which refers to the fact that the original UniFrac definition compares the fraction of edges that connect only tips from one sample via the shortest path rather than edges that connect between samples \citep[Fig.~\ref{fig:unifrac};][]{LozuponeKnightUniFrac05}.
Weighted UniFrac is an abundance weighted version \citep{LozuponeEaWeightedUnifrac07}.
These dissimilarity measures have hundreds of citations.
\citet{evans2012phylogenetic} showed that weighted UniFrac is in fact a specific case of the earth-movers distance, and that the commonly-used randomization procedure for significance estimation has a central limit theorem approximation.
The earth-movers distance between two probability distributions in this case can be defined using a physical analogy as the minimum amount of ``work'' required to move piles of dirt (each dirt pile configuration sized in proportion to one probability distribution) in one configuration to another along the tree, where work is defined as mass times distance.
In this case the size of the dirt piles is proportional to the number of reads mapping to that location in the tree (Fig.~\ref{fig:dirtpiles}).
\citet{chen2012associating} have shown that a partially abundance-weighted variant of UniFrac may have greater power to resolve community differences than either unweighted or weighted UniFrac.
There are clear connections between UniFrac and phylogenetic diversity explored by \cite{faith2009cladistic}, who point out related measures in Faith's earlier work.

@@

There are also other methods with the less ambitious goal of mapping sequences of unknown origin into a so-called fixed \textit{reference tree}, sometimes with uncertainty estimates.
These programs \citep{wu2008simple,monierEaLargeViruses08,vonMeringEaQuantitative08,stark2010mltreemap,matsen2010pplacer,berger2011performance} have various speeds and features.
This work has also spurred development of specialized alignment tools for this mapping process.
\citet{berger2011aligning} focus on the problem of inferring the optimal alignment and insertion of sequences into a tree.
\citet{Mirarab2012-pk} use data set partitioning to improve alignments on subsets of taxa specifically for this application.
\citet{Brown2013-ia} use locality-sensitive hashing to obtain placement more than two orders of magnitude faster than the \textit{pplacer} program of \cite{matsen2010pplacer}.

@@

\noindent
\begin{figure}[ht]
  \caption{\pdLegend}
  \label{fig:pd}
\end{figure}

\noindent
\begin{figure}[ht]
  \caption{\unifracLegend}
  \label{fig:unifrac}
\end{figure}

\noindent
\begin{figure}[ht]
  \caption{\dirtpilesLegend}
  \label{fig:dirtpiles}
\end{figure}

@@

\bibliographystyle{sysbio}
\bibliography{sbreview}
